
##### Assignment for Getting & Cleaning Data : CODEbook_run_analysis.Rmd
<p>CODEbook for run_analysis.R -- describing the variables -- coded by h-rm_tan 22-24_Oct2015</p>
<p>Refer to README_run_analysis.Rmd for details on what the script does.</p>

##### Variables' Descriptions : the variables that run_analysis.R spews out at each stage are as follows :

###### --- (1) READ FILES & ADD VARIABLE NAMES --------------------------------------------------------------

<b>activityL</b> is a 6 x 2 matrix describing the 6 types of activities performed by each participant.
Column1 is numbered 1:6; column2 lists the activities ["LAYING" ; "SITTING" ; "STANDING"; "WALKING"; "WALKING_DOWNSTAIRS" ; "WALKING_UPSTAIRS"]
```{r}
activityL <- read.table("/activity_labels.txt")   # dim(activityL)  # [1] 6 2

activityL
#   V1                 V2
# 1  1            WALKING
# 2  2   WALKING_UPSTAIRS
# 3  3 WALKING_DOWNSTAIRS
# 4  4            SITTING
# 5  5           STANDING
# 6  6             LAYING
```
#####

<b>activityF</b> is a 561 x 2 matrix with 2nd column listing descriptions of the activity features calculated from raw 3-axial angular acceleration and velocity measured by phone's accelerometer and gyroscope. 1st column has rows numbered 1:561.
```{r}
activityF <- read.table("/features.txt")   # dim(activityF)  # [1] 561   2

activityF
#      V1                                   V2
# 1     1                    tBodyAcc-mean()-X
# 2     2                    tBodyAcc-mean()-Y
# 3     3                    tBodyAcc-mean()-Z
# 4     4                     tBodyAcc-std()-X
# 5     5                     tBodyAcc-std()-Y
# 6     6                     tBodyAcc-std()-Z
# 7     7                     tBodyAcc-mad()-X
# 8     8                     tBodyAcc-mad()-Y
# 9     9                     tBodyAcc-mad()-Z
# 10   10                     tBodyAcc-max()-X
# 11   11                     tBodyAcc-max()-Y
# 12   12                     tBodyAcc-max()-Z
# 13   13                     tBodyAcc-min()-X
# 14   14                     tBodyAcc-min()-Y
# 15   15                     tBodyAcc-min()-Z
# 16   16                       tBodyAcc-sma()
# 17   17                  tBodyAcc-energy()-X
# 18   18                  tBodyAcc-energy()-Y
# 19   19                  tBodyAcc-energy()-Z
# 20   20                     tBodyAcc-iqr()-X
# 21   21                     tBodyAcc-iqr()-Y
# 22   22                     tBodyAcc-iqr()-Z
# 23   23                 tBodyAcc-entropy()-X
# 24   24                 tBodyAcc-entropy()-Y
# 25   25                 tBodyAcc-entropy()-Z
# 26   26               tBodyAcc-arCoeff()-X,1
# 27   27               tBodyAcc-arCoeff()-X,2
# 28   28               tBodyAcc-arCoeff()-X,3
# 29   29               tBodyAcc-arCoeff()-X,4
# 30   30               tBodyAcc-arCoeff()-Y,1
# 31   31               tBodyAcc-arCoeff()-Y,2
# 32   32               tBodyAcc-arCoeff()-Y,3
# 33   33               tBodyAcc-arCoeff()-Y,4
# 34   34               tBodyAcc-arCoeff()-Z,1
# 35   35               tBodyAcc-arCoeff()-Z,2
# 36   36               tBodyAcc-arCoeff()-Z,3
# 37   37               tBodyAcc-arCoeff()-Z,4
# ... ... ...
```
#####

<b>FeatureNames</b> is a vector containing string characters derived from the 2nd column of activityF which contains the activity feature descriptions. This vector of descriptions are forced to be a list of "unique" names.
```{r}
## --- make.names for feature variables
FeatureNames <- make.names(activityF[,2], unique = TRUE, allow_ = TRUE)

FeatureNames
#   [1] "tBodyAcc.mean...X"                    "tBodyAcc.mean...Y"                    "tBodyAcc.mean...Z"                   
#   [4] "tBodyAcc.std...X"                     "tBodyAcc.std...Y"                     "tBodyAcc.std...Z"                    
#   [7] "tBodyAcc.mad...X"                     "tBodyAcc.mad...Y"                     "tBodyAcc.mad...Z"                    
#  [10] "tBodyAcc.max...X"                     "tBodyAcc.max...Y"                     "tBodyAcc.max...Z"                    
#  [13] "tBodyAcc.min...X"                     "tBodyAcc.min...Y"                     "tBodyAcc.min...Z"                    
#  [16] "tBodyAcc.sma.."                       "tBodyAcc.energy...X"                  "tBodyAcc.energy...Y"                 
#  [19] "tBodyAcc.energy...Z"                  "tBodyAcc.iqr...X"                     "tBodyAcc.iqr...Y"                    
#  [22] "tBodyAcc.iqr...Z"                     "tBodyAcc.entropy...X"                 "tBodyAcc.entropy...Y"                
#  [25] "tBodyAcc.entropy...Z"                 "tBodyAcc.arCoeff...X.1"               "tBodyAcc.arCoeff...X.2"              
#  [28] "tBodyAcc.arCoeff...X.3"               "tBodyAcc.arCoeff...X.4"               "tBodyAcc.arCoeff...Y.1"              
#  [31] "tBodyAcc.arCoeff...Y.2"               "tBodyAcc.arCoeff...Y.3"               "tBodyAcc.arCoeff...Y.4"              
#  [34] "tBodyAcc.arCoeff...Z.1"               "tBodyAcc.arCoeff...Z.2"               "tBodyAcc.arCoeff...Z.3"              
#  [37] "tBodyAcc.arCoeff...Z.4"               "tBodyAcc.correlation...X.Y"           "tBodyAcc.correlation...X.Z"          
#  [40] "tBodyAcc.correlation...Y.Z"           "tGravityAcc.mean...X"                 "tGravityAcc.mean...Y"                
#  [43] "tGravityAcc.mean...Z"                 "tGravityAcc.std...X"                  "tGravityAcc.std...Y"                 
#  [46] "tGravityAcc.std...Z"                  "tGravityAcc.mad...X"                  "tGravityAcc.mad...Y"                 
#  [49] "tGravityAcc.mad...Z"                  "tGravityAcc.max...X"                  "tGravityAcc.max...Y"                 
#  [52] "tGravityAcc.max...Z"                  "tGravityAcc.min...X"                  "tGravityAcc.min...Y"                 
#  [55] "tGravityAcc.min...Z"                  "tGravityAcc.sma.."                    "tGravityAcc.energy...X"              
#  [58] "tGravityAcc.energy...Y"               "tGravityAcc.energy...Z"               "tGravityAcc.iqr...X"                 
#  ... ... ...               
```
#####

<b>trainX</b> contains 7352 rows of measurements of activity features from the all subjects belonging to the set of TRAINING data, with corresponding activity feature descriptions for 561-column variable names.
```{r}
trainX <- read.table("train/X_train.txt")   # dim(trainX)     # [1] 7352  561
names(trainX) <- FeatureNames 
```
#####

<b>testX</b> contains 2947 rows of measurements of activity features from the all subjects belonging to the set of TEST data, with corresponding activity feature descriptions for 561-column variable names.
```{r}
testX <- read.table("test/X_test.txt")   # dim(testX)      # [1] 2947  561
names(testX) <- FeatureNames 
```
#####

<b>trainY</b> is a 7352-row vector listing corresponding activity coded from 1 to 6 for the types of activity (e.g. "LAYING" ; "SITTING" ; "STANDING"; "WALKING"; "WALKING_DOWNSTAIRS" ; "WALKING_UPSTAIRS" ) performed by subjects belonging to the TRAINING group.
```{r}
trainY <- read.table("train/y_train.txt")   # dim(trainY)     # [1] 7352    1
names(trainY) <- "activity"

unique(trainY)
#     activity
# 1          5
# 28         4
# 52         6
# 79         1
# 126        3
# 151        2
```
#####

<b>testY</b> is a 2947-row vector listing corresponding activity coded from 1 to 6 for the types of activity (e.g. "LAYING" ; "SITTING" ; "STANDING"; "WALKING"; "WALKING_DOWNSTAIRS" ; "WALKING_UPSTAIRS" ) performed by subjects belonging to the TEST group.
```{r}
testY <- read.table("test/y_test.txt")   # dim(testY)      # [1] 2947    1
names(testY) <- "activity"

unique(testY)
#     activity
# 1          5
# 32         4
# 56         6
# 80         1
# 110        3
# 134        2
```
#####

<b>trainSubj</b> is a 7352 x 2 matrix of subject ID codes ranging from 1:30 (inclusive) identifying subjects (subjID) who belong to the TRAINING group (datatype). The rows correspond to those in trainY and trainX. 
```{r}
trainSubj <- read.table("train/subject_train.txt")   # dim(trainSubj)  # [1] 7352    1
names(trainSubj) <- "subjID"
trainSubj <- trainSubj %>% mutate(datatype = "train")   # dim(trainSubj)  # [1] 7352    2

unique(trainSubj)
#      subjID datatype
# 1         1    train
# 348       3    train
# 689       5    train
# 991       6    train
# 1316      7    train
# 1624      8    train
# 1905     11    train
# 2221     14    train
# 2544     15    train
# 2872     16    train
# 3238     17    train
# 3606     19    train
# 3966     21    train
# 4374     22    train
# 4695     23    train
# 5067     25    train
# 5476     26    train
# 5868     27    train
# 6244     28    train
# 6626     29    train
# 6970     30    train
```
#####

<b>testSubj</b> is a 2947 x 2 matrix of subject ID codes ranging from 1:30 (inclusive) identifying subjects (subjID) who belong to the TEST group (datatype). The rows correspond to those in testY and testX. 
```{r}
testSubj <- read.table("test/subject_test.txt")   # dim(testSubj)   # [1] 2947    1
names(testSubj) <- "subjID"
testSubj <- testSubj %>% mutate(datatype = "test")   # dim(testSubj)   # [1] 2947    2   

unique(testSubj)
#      subjID datatype
# 1         2     test
# 303       4     test
# 620       9     test
# 908      10     test
# 1202     12     test
# 1522     13     test
# 1849     18     test
# 2213     20     test
# 2567     24     test
```


###### --- (2) COMBINE & MERGE DATA --------------------------------------------------------------

<b>trainD</b> is a column-wise combination of trainSubj, trainY and trainX to form a 7352 x 564 matrix
```{r}
## --- combineTRAINdata
trainD <- tbl_df(cbind(trainSubj,trainY,trainX))   # Source: local data frame [7,352 x 564]
```

<b>testD</b> is a column-wise combination of testSubj, testY and testX to form a 2947 x 564 matrix
```{r}
## --- combineTRAINdata
testD <- tbl_df(cbind(testSubj,testY,testX))   # Source: local data frame [2,947 x 564]
```

<b>D</b> is a [10,299 x 564 matrix] result of merging trainD and testD by combining (appending) all rows from both tables.
```{r}
## MERGE train&testD
D <- rbind(trainD,testD)   # Source: local data frame [10,299 x 564]
```


###### --- (3) EXTRACT Mean & Std Measures --------------------------------------------------------------

<b>D1</b> is a subset table derived from D that contains only the Activity Feature Measures' Mean and Standard Deviation values for all subjects and activity performed. It has 10299 rows and 89 columns.
```{r}
D1 <- D %>% select(subjID, datatype, activity, contains("mean"), contains("std") ) # dim: [10,299 x 86 +3]
```


The column variable names of D1 are as follows:
```{r}
names(D1)
# [1] "subjID"                               "datatype"                             "activity"
# [4] "tBodyAcc.mean...X"                    "tBodyAcc.mean...Y"                    "tBodyAcc.mean...Z"
# [7] "tGravityAcc.mean...X"                 "tGravityAcc.mean...Y"                 "tGravityAcc.mean...Z"
# [10] "tBodyAccJerk.mean...X"                "tBodyAccJerk.mean...Y"                "tBodyAccJerk.mean...Z"
# [13] "tBodyGyro.mean...X"                   "tBodyGyro.mean...Y"                   "tBodyGyro.mean...Z"
# [16] "tBodyGyroJerk.mean...X"               "tBodyGyroJerk.mean...Y"               "tBodyGyroJerk.mean...Z"
# [19] "tBodyAccMag.mean.."                   "tGravityAccMag.mean.."                "tBodyAccJerkMag.mean.."
# [22] "tBodyGyroMag.mean.."                  "tBodyGyroJerkMag.mean.."              "fBodyAcc.mean...X"
# [25] "fBodyAcc.mean...Y"                    "fBodyAcc.mean...Z"                    "fBodyAcc.meanFreq...X"
# [28] "fBodyAcc.meanFreq...Y"                "fBodyAcc.meanFreq...Z"                "fBodyAccJerk.mean...X"
# [31] "fBodyAccJerk.mean...Y"                "fBodyAccJerk.mean...Z"                "fBodyAccJerk.meanFreq...X"
# [34] "fBodyAccJerk.meanFreq...Y"            "fBodyAccJerk.meanFreq...Z"            "fBodyGyro.mean...X"
# [37] "fBodyGyro.mean...Y"                   "fBodyGyro.mean...Z"                   "fBodyGyro.meanFreq...X"
# [40] "fBodyGyro.meanFreq...Y"               "fBodyGyro.meanFreq...Z"               "fBodyAccMag.mean.."
# [43] "fBodyAccMag.meanFreq.."               "fBodyBodyAccJerkMag.mean.."           "fBodyBodyAccJerkMag.meanFreq.."
# [46] "fBodyBodyGyroMag.mean.."              "fBodyBodyGyroMag.meanFreq.."          "fBodyBodyGyroJerkMag.mean.."
# [49] "fBodyBodyGyroJerkMag.meanFreq.."      "angle.tBodyAccMean.gravity."          "angle.tBodyAccJerkMean..gravityMean."
# [52] "angle.tBodyGyroMean.gravityMean."     "angle.tBodyGyroJerkMean.gravityMean." "angle.X.gravityMean."
# [55] "angle.Y.gravityMean."                 "angle.Z.gravityMean."                 "tBodyAcc.std...X"
# [58] "tBodyAcc.std...Y"                     "tBodyAcc.std...Z"                     "tGravityAcc.std...X"
# [61] "tGravityAcc.std...Y"                  "tGravityAcc.std...Z"                  "tBodyAccJerk.std...X"
# [64] "tBodyAccJerk.std...Y"                 "tBodyAccJerk.std...Z"                 "tBodyGyro.std...X"
# [67] "tBodyGyro.std...Y"                    "tBodyGyro.std...Z"                    "tBodyGyroJerk.std...X"
# [70] "tBodyGyroJerk.std...Y"                "tBodyGyroJerk.std...Z"                "tBodyAccMag.std.."
# [73] "tGravityAccMag.std.."                 "tBodyAccJerkMag.std.."                "tBodyGyroMag.std.."
# [76] "tBodyGyroJerkMag.std.."               "fBodyAcc.std...X"                     "fBodyAcc.std...Y"
# [79] "fBodyAcc.std...Z"                     "fBodyAccJerk.std...X"                 "fBodyAccJerk.std...Y"
# [82] "fBodyAccJerk.std...Z"                 "fBodyGyro.std...X"                    "fBodyGyro.std...Y"
# [85] "fBodyGyro.std...Z"                    "fBodyAccMag.std.."                    "fBodyBodyAccJerkMag.std.."
# [88] "fBodyBodyGyroMag.std.."               "fBodyBodyGyroJerkMag.std.."
```


###### --- (4) USE Descriptive Activity Names --------------------------------------------------------------

After substituting activity codes with activity descriptions, the levels of the variable "activity"" in D1 now reflect the activity descriptions:
```{r}
D1$activity <- activityL[D1$activity,2] #%>% print

levels(D1$activity)
# [1] "LAYING"             "SITTING"            "STANDING"           "WALKING"
# [5] "WALKING_DOWNSTAIRS" "WALKING_UPSTAIRS"
```

###### --- (5) DERIVE Variable Means GROUP by Subj & Activity --------------------------------------------------------------

<b>D2_tidymeans</b> is a 180 x 89 matrix summary of the means of activity feature variables in D1 which are grouped by the activity performed and by each subject (regardless of their group assignment). Each subject has 6 rows of data values; a total of 180rows.
```{r}
## --- MEANS of variables GROUPed by subjID, activity (& datatype) --------------------------------------------------------------
D2_tidymeans <- D1 %>% group_by(subjID, datatype, activity) %>% summarise_each(funs(mean), -subjID:-activity )

# An overview of D2_tidymeans:
D2_tidymeans
# Source: local data frame [180 x 89]
# Groups: subjID, datatype [?]
#
# subjID datatype           activity tBodyAcc.mean...X tBodyAcc.mean...Y tBodyAcc.mean...Z tGravityAcc.mean...X
# (int)    (chr)             (fctr)             (dbl)             (dbl)             (dbl)                (dbl)
# 1       1    train             LAYING         0.2215982      -0.040513953        -0.1132036           -0.2488818
# 2       1    train            SITTING         0.2612376      -0.001308288        -0.1045442            0.8315099
# 3       1    train           STANDING         0.2789176      -0.016137590        -0.1106018            0.9429520
# 4       1    train            WALKING         0.2773308      -0.017383819        -0.1111481            0.9352232
# 5       1    train WALKING_DOWNSTAIRS         0.2891883      -0.009918505        -0.1075662            0.9318744
# 6       1    train   WALKING_UPSTAIRS         0.2554617      -0.023953149        -0.0973020            0.8933511
# 7       2     test             LAYING         0.2813734      -0.018158740        -0.1072456           -0.5097542
# 8       2     test            SITTING         0.2770874      -0.015687994        -0.1092183            0.9404773
# 9       2     test           STANDING         0.2779115      -0.018420827        -0.1059085            0.8969286
# 10      2     test            WALKING         0.2764266      -0.018594920        -0.1055004            0.9130173
# ..    ...      ...                ...               ...               ...               ...                  ...
# Variables not shown: tGravityAcc.mean...Y (dbl), tGravityAcc.mean...Z (dbl), tBodyAccJerk.mean...X (dbl),
# tBodyAccJerk.mean...Y (dbl), tBodyAccJerk.mean...Z (dbl), tBodyGyro.mean...X (dbl), tBodyGyro.mean...Y (dbl),
# tBodyGyro.mean...Z (dbl), tBodyGyroJerk.mean...X (dbl), tBodyGyroJerk.mean...Y (dbl), tBodyGyroJerk.mean...Z (dbl),
# tBodyAccMag.mean.. (dbl), tGravityAccMag.mean.. (dbl), tBodyAccJerkMag.mean.. (dbl), tBodyGyroMag.mean.. (dbl),
# tBodyGyroJerkMag.mean.. (dbl), fBodyAcc.mean...X (dbl), fBodyAcc.mean...Y (dbl), fBodyAcc.mean...Z (dbl),
# fBodyAcc.meanFreq...X (dbl), fBodyAcc.meanFreq...Y (dbl), fBodyAcc.meanFreq...Z (dbl), fBodyAccJerk.mean...X (dbl),
# fBodyAccJerk.mean...Y (dbl), fBodyAccJerk.mean...Z (dbl), fBodyAccJerk.meanFreq...X (dbl), fBodyAccJerk.meanFreq...Y
# (dbl), fBodyAccJerk.meanFreq...Z (dbl), fBodyGyro.mean...X (dbl), fBodyGyro.mean...Y (dbl), fBodyGyro.mean...Z
# (dbl), fBodyGyro.meanFreq...X (dbl), fBodyGyro.meanFreq...Y (dbl), fBodyGyro.meanFreq...Z (dbl), fBodyAccMag.mean..
# (dbl), fBodyAccMag.meanFreq.. (dbl), fBodyBodyAccJerkMag.mean.. (dbl), fBodyBodyAccJerkMag.meanFreq.. (dbl),
# fBodyBodyGyroMag.mean.. (dbl), fBodyBodyGyroMag.meanFreq.. (dbl), fBodyBodyGyroJerkMag.mean.. (dbl),
# fBodyBodyGyroJerkMag.meanFreq.. (dbl), angle.tBodyAccMean.gravity. (dbl), angle.tBodyAccJerkMean..gravityMean.
# (dbl), angle.tBodyGyroMean.gravityMean. (dbl), angle.tBodyGyroJerkMean.gravityMean. (dbl), angle.X.gravityMean.
# (dbl), angle.Y.gravityMean. (dbl), angle.Z.gravityMean. (dbl), tBodyAcc.std...X (dbl), tBodyAcc.std...Y (dbl),
# tBodyAcc.std...Z (dbl), tGravityAcc.std...X (dbl), tGravityAcc.std...Y (dbl), tGravityAcc.std...Z (dbl),
# tBodyAccJerk.std...X (dbl), tBodyAccJerk.std...Y (dbl), tBodyAccJerk.std...Z (dbl), tBodyGyro.std...X (dbl),
# tBodyGyro.std...Y (dbl), tBodyGyro.std...Z (dbl), tBodyGyroJerk.std...X (dbl), tBodyGyroJerk.std...Y (dbl),
# tBodyGyroJerk.std...Z (dbl), tBodyAccMag.std.. (dbl), tGravityAccMag.std.. (dbl), tBodyAccJerkMag.std.. (dbl),
# tBodyGyroMag.std.. (dbl), tBodyGyroJerkMag.std.. (dbl), fBodyAcc.std...X (dbl), fBodyAcc.std...Y (dbl),
# fBodyAcc.std...Z (dbl), fBodyAccJerk.std...X (dbl), fBodyAccJerk.std...Y (dbl), fBodyAccJerk.std...Z (dbl),
# fBodyGyro.std...X (dbl), fBodyGyro.std...Y (dbl), fBodyGyro.std...Z (dbl), fBodyAccMag.std.. (dbl),
# fBodyBodyAccJerkMag.std.. (dbl), fBodyBodyGyroMag.std.. (dbl), fBodyBodyGyroJerkMag.std.. (dbl)
#
```

###### --- (6) WRITE out tidymeans data --------------------------------------------------------------
```{r}
# --- WRITE output D2_tidymeans to working-directory...
# write.table(D2_tidymeans, file = "outputD2_tidymeans.txt", row.name=FALSE, col.names = names(D2_tidymeans), sep = ",")
```
** If you save the file as "outputD2_tidymeans.txt" you can read the file from its directory:
```{r}
outputD2_tidymeans <- read.csv("outputD2_tidymeans.txt")
View(outputD2_tidymeans)
```
outputD2_tidymeans is the same as D2_tidymeans
