
##### Assignment for Getting & Cleaning Data: README_run_analysis.Rmd 
README for run_analysis.R -- describing how the script works: coded by h-rm_tan 22-24_Oct2015.
<br>*Refer to CODEbook_run_analysis.Rmd for description of variables.</p>

##### REFs:
RAW DATA URL: http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones 
<br>Activities measured shown in video: https://www.youtube.com/watch?v=XOEN9W05_4A 

(i) The code should work if you have run_analysis.R and the unzipped "UCI HAR Dataset" data folder present in the working directory of R/Rstudio. Otherwise, paths need to be set, for example (as commented code):

```{r}
## --- SET PATHS --------------------------------------------------------------
# datapath <- "~/Documents/OwnCloud/Coursera_DataScienceTrack_JH/03_Getting&CleaningData/Assignment/UCI HAR Dataset/"
# setwd(datapath)
```

(ii) The required dplr package is loaded...
```{r}
## --- LIBRARIES --------------------------------------------------------------
library(dplyr)
```

###### --- (1) READ FILES & ADD VARIABLE NAMES --------------------------------------------------------------
(1) All the relevant files are being read in 1st by the code as follows:

1.1 First the activity variable description files are being read.
```{r}
## --- variables' descriptions
activityL <- read.table("/activity_labels.txt")   

activityF <- read.table("/features.txt")   
```

1.2 The descriptions of activity features are actually not in a right format for use as variable 'names' so the script resolves this by "making names" from the 2nd column of character strings from activityF. <br>This also allows the measurement data to be appropriately labelled with descriptive variable names. This further enables one to apply the select() function subsequently.  

```{r}
## --- make.names for feature variables
## ** resolving problem with "duplicate-VarNames" for subsequent use of select() --> make.names
FeatureNames <- make.names(activityF[,2], unique = TRUE, allow_ = TRUE)
```

1.3 Next the data with measurements of activity features are read from both training and test datasets. <br>
The activity feature names are also assigned to the corresponding variables for the respective data-tables.

```{r}
## --- measured features -- activityF
trainX <- read.table("train/X_train.txt")   
names(trainX) <- FeatureNames 

testX <- read.table("test/X_test.txt")   
names(testX) <- FeatureNames 
```

1.4 Next the coded activity data corresponding to the measured activity features for both training and test datasets are read and correspondingly assigned with an "activity" variable name.

```{r}
## --- measured activity -- activityL || activityL[trainY$activity,]
trainY <- read.table("train/y_train.txt")   
names(trainY) <- "activity"

testY <- read.table("test/y_test.txt")   
names(testY) <- "activity"
```

1.5 Participant IDs that correspond to the measured activities and activity features are read for both training and test datasets. They are assigned with a variable name: "subjID". 
<br>To faciliate data-handling when both sets of data are combined, an extra variable "datatype" is added using mutate() to help identify if they belong to the training or test dataset.

```{r}
## --- subjs in training & test groups
trainSubj <- read.table("train/subject_train.txt")   
names(trainSubj) <- "subjID"
trainSubj <- trainSubj %>% mutate(datatype = "train")

testSubj <- read.table("test/subject_test.txt")   
names(testSubj) <- "subjID"
testSubj <- testSubj %>% mutate(datatype = "test")
```

###### --- (2) COMBINE & MERGE DATA --------------------------------------------------------------
(2) Now the corresponding parts of training and test datasets can be combined in 2 steps.

2.1 First -- the script combines the feature measurements for corresponding activity performed by each participant belonging to either the training or test group. This is achieved using cbind() to combine the relevant data variables by "columns".

```{r}
## --- COMBINE components of train DATA & test DATA --------------------------------------------------------------
## --- combineTRAINdata
trainD <- tbl_df(cbind(trainSubj,trainY,trainX))

## --- combineTRAINdata
testD <- tbl_df(cbind(testSubj,testY,testX))
```

2.2 Second -- the set of trainingData and testData are subsequently merged by "rows" using rbind() to create a single dataset which already has the appropriate variable name labels (see 1.2 above).

```{r}
## --- MERGE DATA --------------------------------------------------------------
## MERGE train&testD
D <- rbind(trainD,testD)
```

###### --- (optional) Free Up Working Memory --------------------------------------------------------------
Now that both datasets are merged, we could free up some working memory by clearing all the training and test variables, where appropriate. However, this is currently commented out in the code. 
```{r}
# # --- clear workingMEM...
# rm(list = c( ls(pattern = "^test"), ls(pattern = "^train") ) )
```

###### --- (3) EXTRACT Mean & Std Measures --------------------------------------------------------------
(3) Next the script selects only measured mean and standard deviation from the activity features variables. 
<br>The subjID, datatype and activity information is kept while selecting only variable names that contain either "mean" or "std" using a chained process of select(). The output is assigned as "D1".

```{r}
D1 <- D %>% select(subjID, datatype, activity, contains("mean"), contains("std") ) 
```

###### --- (4) USE Descriptive Activity Names --------------------------------------------------------------
(4) Subsequently, the exisiting activity codes [1:6] are replaced with their corresponding activity descriptions: "LAYING" ; "SITTING" ; "STANDING"; "WALKING"; "WALKING_DOWNSTAIRS" ; "WALKING_UPSTAIRS". 
<br>This is achieved by using the values contained within the current factor "activity" in D1, which ranges from 1 to 6, to index  "activityL", whose 2nd column is a vector of activity labels, and then substituting this for the "activity" factor in D1 in a single step.

```{r}
D1$activity <- activityL[D1$activity,2] #%>% print
```

###### --- (5) DERIVE Variable Means GROUPED by Subj & Activity --------------------------------------------------------------
(5) Next the script creates a 2nd independent tidy data set from D1, which is assigned as "D2_tidymeans". 
<br>D2_tidymeans contains information on the average of each of the activity feature variable for each activity and each subject.

```{r}
## --- MEANS of variables GROUPed by subjID, activity (& datatype) --------------------------------------------------------------
D2_tidymeans <- D1 %>% group_by(subjID, datatype, activity) %>% summarise_each(funs(mean), -subjID:-activity )
```

###### --- (6) WRITE out tidymeans data --------------------------------------------------------------
(6) Finally the output D2_tidymeans is written out as a comma-separated text file "outputD2_tidymeans.txt", which could be read with read.csv() and viewed in R/Rstudio.

```{r}
# --- WRITE output D2_tidymeans to working-directory...
write.table(D2_tidymeans, file = "outputD2_tidymeans.txt", row.name=FALSE, col.names = names(D2_tidymeans), sep = ",")
```

